---
title: "Saccharide profile analysis"
author: "Daniel Garrido-Sanz (daniel.garridosanz@unil.ch)"
affiliation: "Kell Lab | Department of fundamental microbiology | University of Lausanne"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::html_clean:
    code_folding: hide
    toc: true
    thumbnails: false
    toc_depth: 3
pkgdown:
  as_is: true
---

We have selected 5 strains to characterize their ability to use 2 monosacharides (glucose and fructose) and four disacharadies (sucrose, maltose, trehalose and cellobiose). In addition, we have characterized the ability of growin with co-cultured with a Pantoea.

```{r}
library(tidyverse)
library(ggplot2)
library(ggpubr)
```

#0. Load data

```{r}
cfus = read.table("Saccharide_profile_CFUs.txt", header = T, sep = "\t", quote = "")
head(cfus)
```

#1. Check controls.

Check if any controls have growth and remove it not.

```{r}
if(count(cfus %>% filter(strain_name == "Control" & CFUs_mL != 0)) == 0){
  print(cfus %>% filter(strain_name == "Control" & CFUs_mL != 0))
  cfus = cfus %>% filter(strain_name != "Control")
}else{
  print(cfus %>% filter(strain_name == "Control" & CFUs_mL != 0))
}

head(cfus)
```

#2. Inoculum

Extract CFUs in M9 base media (without carbon sources) and calculate mean values per strain

```{r}
#plot distribution of inoculum
ggplot(cfus %>% filter(carbon_sour == "M9_NoCarbon"), aes(x=strain_name, y=log10(CFUs_mL)))+
  geom_boxplot()+
  geom_jitter() +
  ggtitle("Inoculum size")

inoculum = cfus %>% filter(carbon_sour == "M9_NoCarbon") %>% 
  group_by(strain_name) %>% 
  summarize(CFUs.mL_M9 = mean(CFUs_mL))
head(inoculum)

```

#3. Normalize growth

Now we can normalize the growth of each strain by subtracting the M9 base media CFUs.

```{r}
cfus = cfus %>% left_join(inoculum, by="strain_name") %>%
  mutate(CFUs.mL_growth = CFUs_mL - CFUs.mL_M9)
head(cfus)
```

#4. Plot


```{r}
# Combine the datasets
cfus_combined = cfus %>%
  filter(carbon_sour != "M9_NoCarbon") %>%
  mutate(Experiment_Label = case_when(
    Experiment == "monoculture" ~ "Monoculture",
    Experiment == "coculture_SbRB3" ~ "SbRB3 Coculture",
    Experiment == "coculture_DGS31" ~ "DGS31 Coculture")) %>%
  group_by(sample_name, strain_name, carbon_sour, Experiment_Label) %>%
  mutate(mean_CFUs = mean(CFUs.mL_growth, na.rm = TRUE)) %>%
  ungroup()

# Create the plot
ggplot(cfus_combined,
       aes(
         x = ifelse(Experiment == "monoculture", strain_name,
                    paste0(gsub("_(.*)", "", sample_name), "-", strain_name)),
         y = log10(CFUs.mL_growth),
         fill = carbon_sour)) +
  geom_col(position = "identity", aes(y=log10(mean_CFUs))) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(shape = 16, width = 0.2) +
  coord_cartesian(ylim = c(3, 10)) +
  scale_fill_manual(values = c("#1b9e77", "#66a61e", "#7570b3", "#e7298a", "#d95f02", "#e6ab02")) +
  facet_grid(carbon_sour ~ Experiment_Label, scales = "free_x", space = "free_x") +
  ylab("log10(Norm. CFUs/mL)")+ xlab("Strain")+
  theme_bw()+
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    panel.spacing = unit(1, "lines"))
```
